---
title: "comparing cities"
output: html_document
date: "2023-08-04"
---

# Setting up

```{r}
library(tidyverse)
library(readxl)
library(sf)
library(ggplot2)
library(reshape2)
library(stringr)
library(hrbrthemes)
library(gridExtra)
```

# Reading in the data

```{r}
cluster <- read.csv("data/cluster_results_prop.csv")
entropy <- read.csv("data/entropy_results_mean.csv")
nmi <- read.csv("data/nmi_results.csv")
fua_geom <- st_read("fua.gpkg")
fua <- fua_geom %>% st_drop_geometry()
uc_gdp <- read.csv("data/uc_gdp.csv")
shannon <- read.csv("data/shannon_entropy_output.csv")
```

## cleaning the data

clean fua data to merge with FUA_data

```{r}
# select the columns to keep
fua_cleaned <- fua %>% select(UC_num, eFUA_name, Commuting, Cntry_ISO, FUA_area, UC_area, FUA_p_2015, UC_p_2015, Com_p_2015)

# convert city names to all lowercase and change space to underscore and other characters
fua_cleaned$eFUA_name <- tolower(fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub(" ", "_", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("í", "i", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("é", "e", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("ç", "c", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("-", "_", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("â", "a", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("ú", "u", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("á", "a", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("ó", "o", fua_cleaned$eFUA_name)
fua_cleaned$eFUA_name <- gsub("ã", "a", fua_cleaned$eFUA_name)
  
# create city_name column
fua_cleaned$city_name <- paste0(fua_cleaned$Cntry_ISO, "_", fua_cleaned$eFUA_name)

# remove ISO and Cities column and then bring city_name to front 
fua_cleaned <- subset(fua_cleaned, select = -c(Cntry_ISO, eFUA_name))
```

solve the issue for multiple cities with separate rows

```{r}
fua_cleaned_2 <- fua_cleaned %>%
  relocate(city_name) %>%
  group_by(city_name) %>%
  summarise(UC_num = sum(UC_num, na.rm = TRUE),
            FUA_area  = sum(FUA_area, na.rm = TRUE), 
            UC_area = sum(UC_area, na.rm = TRUE), 
            FUA_p_2015 = sum(FUA_p_2015, na.rm = TRUE), 
            UC_p_2015 = sum(UC_p_2015, na.rm = TRUE), 
            Com_p_2015 = sum(Com_p_2015, na.rm = TRUE))
```

create df for merging

```{r}
# join with cities column
fua_cleaned_2 <- fua_cleaned_2 %>% 
  mutate(FUA_pop_density = FUA_p_2015/FUA_area,
         UC_pop_density = UC_p_2015/UC_area,
         UC_FUA_pop_ratio = UC_p_2015/FUA_p_2015,
         UC_FUA_area_ratio = UC_area/FUA_area,
         UC_FUA_pop_density_ratio = UC_pop_density/FUA_pop_density) %>% 
  select(-Com_p_2015)
```

make mergable by column name

```{r}
colnames(shannon)[1] <- "city_name"
```


merge the dataframes

```{r}
results <- fua_cleaned_2 %>% 
  left_join(cluster, by = "city_name") %>%
  left_join(entropy, by = "city_name") %>% 
  left_join(nmi, by = "city_name") %>%
  left_join(shannon, by = "city_name") %>%
  left_join(uc_gdp, by = "city_name")

# remove the rows with no data 
results <- results %>%
  filter(!is.na(AM_per))

# convert city_column to row index
results <- results %>% 
  remove_rownames %>% 
  column_to_rownames(var="city_name")
```

# Analysis 

add additional columns 

```{r}
results$log_pop <- log(results$FUA_p_2015)
results$log_gdp <- log(results$gdp_2015_sum)
results$gdp_per_capita <- (results$gdp_2015_sum)/(results$UC_p_2015)
```


## Shannon Entropy - absolute

```{r}
# subset to columns
results_entropyS <- results %>%
  select(FUA_p_2015, UC_FUA_pop_density_ratio, gdp_2015_sum, log_pop, log_gdp, gdp_per_capita, FUA_pop_density, Shannon_Absolute_AM, Shannon_Absolute_BH, Shannon_Absolute_NL, Shannon_Absolute_PE, Shannon_Absolute_SI, Shannon_Absolute_WP, )
```

```{r}
# melt data
entropyS <- melt(results_entropyS, id.vars = c("FUA_p_2015", "UC_FUA_pop_density_ratio", "gdp_2015_sum", "log_pop", "log_gdp", "gdp_per_capita", "FUA_pop_density"))
```

```{r}
# clean data - replace data source name, identify value
entropyS$data_source <- str_sub(entropyS$variable, -2, -1)
entropyS$shannon_entropy <- entropyS$value

# drop unnecessary columns
entropyS <- entropyS %>% select(-c(value, variable))
```

### visualise

```{r, fig.height=10, fig.width=15}
ggplot(entropyS, aes(x = log_pop, y = shannon_entropy)) + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="log(population)", y = "Shannon's Entropy") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 4))
```

```{r, fig.height=10, fig.width=15}
ggplot(entropyS, aes(x = gdp_per_capita, y = shannon_entropy)) + geom_point(size = 1) + geom_smooth(method=lm, color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="GDP per capita", y = "Shannon's Entropy") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 4))
```

## Shannon entropy - relative

```{r}
# subset to columns
results_entropyS_2 <- results %>%
  select(FUA_p_2015, UC_FUA_pop_density_ratio, gdp_2015_sum, log_pop, log_gdp, gdp_per_capita, Shannon_Relative_AM, Shannon_Relative_BH, Shannon_Relative_NL, Shannon_Relative_PE, Shannon_Relative_SI, Shannon_Relative_WP)
```

```{r}
# melt data
entropyS_2 <- melt(results_entropyS_2, id.vars = c("FUA_p_2015", "UC_FUA_pop_density_ratio", "gdp_2015_sum", "log_pop", "log_gdp", "gdp_per_capita"))
```

```{r}
# clean data - replace data source name, identify value
entropyS_2$data_source <- str_sub(entropyS_2$variable, -2, -1)
entropyS_2$shannon_entropy <- entropyS_2$value

# drop unnecessary columns
entropyS_2 <- entropyS_2 %>% select(-c(value, variable))
```

```{r, fig.height=10, fig.width=15}
ggplot(entropyS_2, aes(x = log_pop, y = shannon_entropy)) + geom_point(size = 1) + geom_smooth(method=lm, color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="log(population)", y = "Relative Shannon's Entropy") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 1))
```

```{r, fig.height=10, fig.width=15}
ggplot(entropyS_2, aes(x = gdp_per_capita, y = shannon_entropy)) + geom_point(size = 1) + geom_smooth(method=lm, color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="GDP per capita", y = "Relative Shannon's Entropy") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 1))
```

## Batty Entropy 

```{r}
# subset to columns
results_entropyB <- results %>%
  select(FUA_p_2015, UC_FUA_pop_density_ratio, log_pop, log_gdp, gdp_2015_sum,gdp_per_capita, entropy_AM_50, entropy_BH_50, entropy_NL_50, entropy_PE_50, entropy_SI_50, entropy_WP_50)
```

```{r}
# melt data
entropyB <- melt(results_entropyB, id.vars = c("FUA_p_2015", "UC_FUA_pop_density_ratio", "gdp_2015_sum", "log_pop", "log_gdp", "gdp_per_capita"))
```

```{r}
# clean data - replace data source name, identify value
entropyB$data_source <- str_sub(entropyB$variable, 9, 10)
entropyB$batty_entropy <- entropyB$value

# drop unnecessary columns
entropyB <- entropyB %>% select(-c(value, variable))
```

### visualisation

```{r, fig.height=10, fig.width=15}
ggplot(entropyB, aes(x = log_pop, y = batty_entropy)) + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="log(population)", y = "Relative Batty's Entropy") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0.84, 0.96))

```


```{r, fig.height=10, fig.width=15}
ggplot(entropyB, aes(x = gdp_per_capita, y = batty_entropy))  + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="GDP per capita", y = "Relative Batty's Entropy") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0.84, 0.96))
```

```{r, fig.height=10, fig.width=15}
ggplot(entropyB, aes(x = UC_FUA_pop_density_ratio, y = batty_entropy))  + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="Population density ratio (UC/FUA)", y = "NMI with Predicted Employment") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0.84, 0.96))
```

## NMI

```{r}
# subset to columns
results_nmiPE <- results %>%
  select(FUA_p_2015, UC_FUA_pop_density_ratio, gdp_2015_sum, log_pop, log_gdp, nmi.AM.PE, nmi.BH.PE, nmi.NL.PE, nmi.PE.SI, nmi.PE.WP, gdp_per_capita)
```

```{r}
# melt data
nmiPE <- melt(results_nmiPE, id.vars = c("FUA_p_2015", "UC_FUA_pop_density_ratio", "gdp_2015_sum", "log_pop", "log_gdp", "gdp_per_capita"))
```

```{r}
# clean data - replace data source name, identify value
nmiPE <- nmiPE %>% 
  mutate(nmi_PE = value,
         data_source = if_else(variable == "nmi.AM.PE"|variable == "nmi.BH.PE"|variable == "nmi.NL.PE", str_sub(variable, -5, -4), str_sub(variable, -2, -1)))

# drop unnecessary columns
nmiPE <- nmiPE %>% select(-c(value, variable))
```

### visualisation

```{r, fig.height=10, fig.width=15}
ggplot(nmiPE, aes(x = log_pop, y = nmi_PE)) + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="log(population)", y = "NMI with Predicted Employment") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 1.0))
```


```{r, fig.height=10, fig.width=15}
ggplot(nmiPE, aes(x = gdp_per_capita, y = nmi_PE)) + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="log(gdp)", y = "NMI with Predicted Employment") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 1.0))
```

```{r, fig.height=10, fig.width=15}
ggplot(nmiPE, aes(x = UC_FUA_pop_density_ratio, y = nmi_PE)) + geom_point(size = 1) + geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) + facet_wrap(~data_source, scales = "free")+ labs(x ="Population density ratio (UC/FUA)", y = "NMI with Predicted Employment") + theme_classic(base_size = 20) + scale_y_continuous(limits = c(0, 1.0)) 
```

# Regression model

```{r}
lm_entropyS_pop_AM_2 <- lm(Shannon_Relative_AM ~ log_pop, data = results_entropyS_2)
summary(lm_entropyS_pop_AM_2)
```

```{r}
lm_entropyS_pop_BH_2 <- lm(Shannon_Relative_BH ~ log_pop, data = results_entropyS_2)
summary(lm_entropyS_pop_BH_2)
```

```{r}
lm_entropyS_pop_NL_2 <- lm(Shannon_Relative_NL ~ log_pop, data = results_entropyS_2)
summary(lm_entropyS_pop_NL_2)
```

```{r}
lm_entropyS_pop_PE_2 <- lm(Shannon_Relative_PE ~ log_pop, data = results_entropyS_2)
summary(lm_entropyS_pop_PE_2)
```

```{r}
lm_entropyS_pop_SI_2 <- lm(Shannon_Relative_SI ~ log_pop, data = results_entropyS_2)
summary(lm_entropyS_pop_SI_2)
```

```{r}
lm_entropyS_pop_WP_2 <- lm(Shannon_Relative_WP ~ log_pop, data = results_entropyS_2)
summary(lm_entropyS_pop_WP_2)
```
```{r}
lm_entropyS_pop_AM_2 <- lm(Shannon_Relative_AM ~ gdp_per_capita, data = results_entropyS_2)
summary(lm_entropyS_pop_AM_2)
```

```{r}
lm_entropyS_pop_BH_2 <- lm(Shannon_Relative_BH ~ gdp_per_capita, data = results_entropyS_2)
summary(lm_entropyS_pop_BH_2)
```

```{r}
lm_entropyS_pop_NL_2 <- lm(Shannon_Relative_NL ~ gdp_per_capita, data = results_entropyS_2)
summary(lm_entropyS_pop_NL_2)
```

```{r}
lm_entropyS_pop_PE_2 <- lm(Shannon_Relative_PE ~ gdp_per_capita, data = results_entropyS_2)
summary(lm_entropyS_pop_PE_2)
```

```{r}
lm_entropyS_pop_SI_2 <- lm(Shannon_Relative_SI ~ gdp_per_capita, data = results_entropyS_2)
summary(lm_entropyS_pop_SI_2)
```

```{r}
lm_entropyS_pop_WP_2 <- lm(Shannon_Relative_WP ~ gdp_per_capita, data = results_entropyS_2)
summary(lm_entropyS_pop_WP_2)
```

## Shannon's entropy

```{r}
lm_entropyS_pop_AM <- lm(Shannon_Absolute_AM ~ log_pop, data = results_entropyS)
summary(lm_entropyS_pop_AM)
```

```{r}
lm_entropyS_pop_BH <- lm(Shannon_Absolute_BH ~ log_pop, data = results_entropyS)
summary(lm_entropyS_pop_BH)
```

```{r}
lm_entropyS_pop_NL <- lm(Shannon_Absolute_NL ~ log_pop, data = results_entropyS)
summary(lm_entropyS_pop_NL)
```

```{r}
lm_entropyS_pop_PE <- lm(Shannon_Absolute_PE ~ log_pop, data = results_entropyS)
summary(lm_entropyS_pop_PE)
```

```{r}
lm_entropyS_pop_SI <- lm(Shannon_Absolute_SI ~ log_pop, data = results_entropyS)
summary(lm_entropyS_pop_SI)
```

```{r}
lm_entropyS_pop_WP <- lm(Shannon_Absolute_WP ~ log_pop, data = results_entropyS)
summary(lm_entropyS_pop_WP)
```

```{r}
lm_entropyS_gdp_AM <- lm(Shannon_Absolute_AM ~ log_pop + gdp_per_capita, data = results_entropyS)
summary(lm_entropyS_gdp_AM)
```

```{r}
lm_entropyS_gdp_BH <- lm(Shannon_Absolute_BH~log_pop + gdp_per_capita, data = results_entropyS)
summary(lm_entropyS_gdp_BH)
```

```{r}
lm_entropyS_gdp_NL <- lm(Shannon_Absolute_NL~log_pop + gdp_per_capita, data = results_entropyS)
summary(lm_entropyS_gdp_NL)
```

```{r}
lm_entropyS_gdp_PE <- lm(Shannon_Absolute_PE~ log_pop + gdp_per_capita, data = results_entropyS)
summary(lm_entropyS_gdp_PE)
```

```{r}
lm_entropyS_gdp_SI <- lm(Shannon_Absolute_SI~log_pop + gdp_per_capita, data = results_entropyS)
summary(lm_entropyS_gdp_SI)
```

```{r}
lm_entropyS_gdp_WP <- lm(Shannon_Absolute_WP ~ log_pop + gdp_per_capita, data = results_entropyS)
summary(lm_entropyS_gdp_WP)
```

## Batty's entropy

```{r}
lm_entropyB_pop_AM <- lm(entropy_AM_50 ~ log_pop, data = results_entropyB)
summary(lm_entropyB_pop_AM)
```

```{r}
lm_entropyB_pop_BH <- lm(entropy_BH_50 ~ log_pop, data = results_entropyB)
summary(lm_entropyB_pop_BH)
```

```{r}
lm_entropyB_pop_NL <- lm(entropy_NL_50 ~ log_pop, data = results_entropyB)
summary(lm_entropyB_pop_NL)
```

```{r}
lm_entropyB_pop_PE <- lm(entropy_PE_50 ~ log_pop, data = results_entropyB)
summary(lm_entropyB_pop_PE)
```

```{r}
lm_entropyB_pop_SI <- lm(entropy_SI_50 ~ log_pop, data = results_entropyB)
summary(lm_entropyB_pop_SI)
```

```{r}
lm_entropyB_pop_WP <- lm(entropy_WP_50 ~ log_pop, data = results_entropyB)
summary(lm_entropyB_pop_WP)
```

```{r}
lm_entropyB_gdp_AM <- lm(entropy_AM_50 ~ gdp_per_capita + log_pop, data = results_entropyB)
summary(lm_entropyB_gdp_AM)
```

```{r}
lm_entropyB_gdp_BH <- lm(entropy_BH_50 ~ log_pop + gdp_per_capita, data = results_entropyB)
summary(lm_entropyB_gdp_BH)
```

```{r}
lm_entropyB_gdp_NL <- lm(entropy_NL_50 ~ log_pop + gdp_per_capita, data = results_entropyB)
summary(lm_entropyB_gdp_NL)
```

```{r}
lm_entropyB_gdp_PE <- lm(entropy_PE_50 ~ log_pop + gdp_per_capita, data = results_entropyB)
summary(lm_entropyB_gdp_PE)
```

```{r}
lm_entropyB_gdp_SI <- lm(entropy_SI_50 ~ log_pop + gdp_per_capita, data = results_entropyB)
summary(lm_entropyB_gdp_SI)
```

```{r}
lm_entropyB_gdp_WP <- lm(entropy_WP_50 ~ log_pop + gdp_per_capita, data = results_entropyB)
summary(lm_entropyB_gdp_WP)
```

## NMI

```{r}
lm_nmi_pop_AM <- lm(nmi.AM.PE ~ log_pop, data = results_nmiPE)
summary(lm_nmi_pop_AM)
```

```{r}
lm_nmi_pop_BH <- lm(nmi.BH.PE ~ log_pop, data = results_nmiPE)
summary(lm_nmi_pop_BH)
```

```{r}
lm_nmi_pop_NL <- lm(nmi.NL.PE ~ log_pop, data = results_nmiPE)
summary(lm_nmi_pop_NL)
```

```{r}
lm_nmi_pop_SI <- lm(nmi.PE.SI ~ I(1/log_pop), data = results_nmiPE)
summary(lm_nmi_pop_SI)
```

```{r}
lm_nmi_pop_WP <- lm(nmi.PE.WP ~ I(1/log_pop), data = results_nmiPE)
summary(lm_nmi_pop_WP)
```

```{r}
lm_nmi_gdp_AM <- lm(nmi.AM.PE ~ log_gdp, data = results_nmiPE)
summary(lm_nmi_gdp_AM)
```

```{r}
lm_nmi_gdp_BH <- lm(nmi.BH.PE ~ log_gdp, data = results_nmiPE)
summary(lm_nmi_gdp_BH)
```

```{r}
lm_nmi_gdp_NL <- lm(nmi.NL.PE ~ log_gdp, data = results_nmiPE)
summary(lm_nmi_gdp_NL)
```

```{r}
lm_nmi_gdp_SI <- lm(nmi.PE.SI ~ log_gdp, data = results_nmiPE)
summary(lm_nmi_gdp_SI)
```

```{r}
lm_nmi_gdp_WP <- lm(nmi.PE.WP ~ log_gdp, data = results_nmiPE)
summary(lm_nmi_gdp_WP)
```

```{r}
cor.test(results_nmiPE$log_pop, results$gdp_per_capita)
```

# Additional Figures

```{r}
AM <- ggplot(results, aes(x = Shannon_Absolute_AM, y = entropy_AM_50)) + geom_point(size = 1) + xlab("Shannon's entropy") + ylab("Batty's spatial entropy (relative)") + ggtitle ("OSM amenities (AM)")
AM
```

```{r}
BH <- ggplot(results, aes(x = Shannon_Absolute_BH, y = entropy_BH_50)) + geom_point(size = 1) + xlab("Shannon's entropy") + ylab("Batty's spatial entropy (relative)") + ggtitle ("Building Height (BH)")
BH
```

```{r}
NL <- ggplot(results, aes(x = Shannon_Absolute_NL, y = entropy_NL_50)) + geom_point(size = 1) + xlab("Shannon's entropy") + ylab("Batty's spatial entropy (relative)") + ggtitle ("Night Light (NL)")
NL
```

```{r}
PE <- ggplot(results, aes(x = Shannon_Absolute_PE, y = entropy_PE_50)) + geom_point(size = 1) + xlab("Shannon's entropy") + ylab("Batty's spatial entropy (relative)") + ggtitle ("Predicted Employment (PE)")
PE
```

```{r}
SI <- ggplot(results, aes(x = Shannon_Absolute_SI, y = entropy_SI_50)) + geom_point(size = 1) + xlab("Shannon's entropy") + ylab("Batty's spatial entropy (relative)") + ggtitle ("OSM street intersections (SI)")
SI
```

```{r}
WP <- ggplot(results, aes(x = Shannon_Absolute_WP, y = entropy_WP_50)) + geom_point(size = 1) + xlab("Shannon's entropy") + ylab("Batty's spatial entropy (relative)") + ggtitle ("Population (WP)")
WP
```

```{r, fig.height = 10, fig.width=15}
grid.arrange(AM, BH, NL, PE, SI, WP, nrow = 2)
```

