---
title: "Entropy"
output: html_document
date: "2023-07-11"
---

# Setting up 

```{r, message=FALSE, warning=FALSE}
library(sf)
library(dplyr)
library(SpatEntropy)
library(maptools)
```

# Reading in data

setting up procedure for reading in cluster and changing cluster object to sp

```{r}
# basic procedure 
input_path <- "data/cluster/"
files <- list.files(input_path, pattern = "[.]gpkg$", full.names=T)
for (f in files){
  # read in file
  shp <- st_read(f) 
  shp <- shp %>% st_transform(3857)
  # calculate centroid 
  shp_cent <- st_centroid(shp)
  # filter to grids that are classified as hotspots
  shp_AM <- shp_cent %>% 
    filter(., econ_cluster_AM == 1) 
  shp_BH <- shp_cent %>% 
    filter(., econ_cluster_BH == 1) 
  shp_NL <- shp_cent %>% 
    filter(., econ_cluster_NL == 1) 
  shp_PE <- shp_cent %>% 
    filter(., econ_cluster_PE == 1) 
  shp_SI <- shp_cent %>% 
    filter(., econ_cluster_SI == 1) 
  shp_WP <- shp_cent %>% 
    filter(., econ_cluster_WP == 1) 
  # Convert to sp objects
  shp_AM <- as_Spatial(shp_AM)
  shp_BH <- as_Spatial(shp_BH)
  shp_NL <- as_Spatial(shp_NL)
  shp_PE <- as_Spatial(shp_PE)
  shp_SI <- as_Spatial(shp_SI)
  shp_WP <- as_Spatial(shp_WP)
  # get name to save
  shp_name <- gsub(".*data/cluster/(.+)_cluster.gpkg.*", "\\1", f)
  shp_name_AM <- paste0(shp_name, "_AM")
  shp_name_BH <- paste0(shp_name, "_BH")
  shp_name_NL <- paste0(shp_name, "_NL")
  shp_name_PE <- paste0(shp_name, "_PE")
  shp_name_SI <- paste0(shp_name, "_SI")
  shp_name_WP <- paste0(shp_name, "_WP")
  # save with new name
  assign(shp_name_AM, shp_AM, envir = .GlobalEnv)
  assign(shp_name_BH, shp_BH, envir = .GlobalEnv)
  assign(shp_name_NL, shp_NL, envir = .GlobalEnv)
  assign(shp_name_PE, shp_PE, envir = .GlobalEnv)
  assign(shp_name_SI, shp_SI, envir = .GlobalEnv)
  assign(shp_name_WP, shp_WP, envir = .GlobalEnv)
}
# remove the temporary objects
rm(shp)
rm(shp_cent)
rm(shp_AM)
rm(shp_BH)
rm(shp_NL)
rm(shp_PE)
rm(shp_SI)
rm(shp_WP)
```

Read in the boundaries for cities and convert into sp objects

```{r}
# basic procedure
input_path_boundaries <- "data/FUA/"
files_boundaries <- list.files(input_path_boundaries, pattern = "[.]gpkg$", full.names=T)
for (f in files_boundaries){
  # read in file
  shp <- st_read(f)
  shp <- shp %>% st_transform(3857)
  shp <- st_buffer(shp, 500)
  # Convert to sp objects
  shp_sp <- as_Spatial(shp)
  # get name to save
  shp_name <- gsub(".*data/FUA/(.+)_fua.gpkg", "\\1", f)
  # save with new name
  assign(paste(shp_name), shp_sp, envir = .GlobalEnv)
}
# remove the temporary objects
rm(shp)
rm(shp_sp)
```


# Analysis

### Create ppp object

Coerce from SpatialPolygons to an object of class “owin” (observation window)

```{r}
# basic procedure
spolygondfs <- Filter(function(x) is(x, "SpatialPolygonsDataFrame"), mget(ls()))

for(i in 1:length(spolygondfs)) {
  owin_city <- as.owin(spolygondfs[[i]])
  owin_name <- paste0(names(spolygondfs)[i], "_owin") 
  assign(owin_name, owin_city, envir = .GlobalEnv)
}
# remove temporary files
rm(owin_city)
```

Extract coordinates from SpatialPointsDataFrame and create a ‘ppp’ (point pattern) object

```{r}
# basic procedure
spointdfs <- Filter(function(x) is(x, "SpatialPointsDataFrame"), mget(ls()))

for (i in 1:length(spointdfs)){
  pts <- coordinates(spointdfs[[i]])
  # specify pattern of owin 
  city_pattern <- gsub("(.+)_(AM|BH|NL|PE|SI|WP).*", "\\1", names(spointdfs)[i])
  # select owin object
  list_df_owin <- Filter(function(x) is(x, "owin"), mget(ls(pattern = city_pattern)))
  df_owin <- list_df_owin[[1]]
  # create ppp object
  p <- ppp(pts[,1], pts[,2], window = df_owin)
  ppp_name <- paste0(names(spointdfs)[i], "_ppp") 
  assign(ppp_name, p, envir = .GlobalEnv)
}
# remove temporary files
rm(pts)
rm(p)
rm(list_df_owin)
rm(df_owin)
```

calculation of spatial entropy (Batty)

```{r}
# basic procedure
ppps <- Filter(function(x) is(x, "ppp"), mget(ls()))
```

```{r}
# function for producing the relative spatial entropy 
rel_batty <- function(cluster_ppp, part_n){
  batty_ent = batty(cluster_ppp, partition = part_n)
  return (batty_ent$rel.batty)
  }
```

```{r}
# partition = 10
for (i in 1:length(ppps)){
  batty_list <- replicate(100, rel_batty(ppps[[i]], 10), simplify=FALSE)
  batty_unlist <- as.numeric(unlist(batty_list))
  list_name <- paste0(names(ppps)[i], "_10")
  assign(list_name, batty_unlist, envir = .GlobalEnv)
}
```

```{r}
# partition = 20
for (i in 1:length(ppps)){
  batty_list <- replicate(100, rel_batty(ppps[[i]] , 20), simplify=FALSE)
  batty_unlist <- as.numeric(unlist(batty_list))
  list_name <- paste0(names(ppps)[i], "_20")
  assign(list_name, batty_unlist, envir = .GlobalEnv)
}
```

```{r}
# partition = 50
for (i in 1:length(ppps)){
  batty_list <- replicate(100, rel_batty(ppps[[i]], 50), simplify=FALSE)
  batty_unlist <- as.numeric(unlist(batty_list))
  list_name <- paste0(names(ppps)[i], "_50")
  assign(list_name, batty_unlist, envir = .GlobalEnv)
}
```

```{r}
# store the results as a 
```


# Output 

```{r}
## get all objects in environment 
all_objects <- ls()

# create a list to store numeric objects
numeric_objects <- list()

# iterate through all objects and check if they are numeric
for (name in all_objects){
  if (is.numeric(get(name))) {
    numeric_objects[[name]] <- get(name)
  }
}
```


```{r}
# set column names and convert the list of numeric objects into a dataframe
col_names <- names(numeric_objects)
entropy_df <- data.frame(numeric_objects)
colnames(entropy_df) <- col_names
```

```{r}
# save as csv
write.csv(entropy_df, "data/entropy_results.csv", row.names = TRUE)
```

