---
title: "Moran's I"
output: html_document
date: "2023-07-18"
---

# Setting up

```{r}
library(sf)
library(spdep)
library(dplyr)
library(tidyr)
library(purrr)
library(tmap)
```

# Reading in the data

- predicted employment 

```{r}
# basic procedure 
input_path_pe <- "data/relative/"
files_pe <- list.files(input_path_pe, pattern = "[.]gpkg$", full.names=T)
for (f in files_pe){
  # read in file
  shp <- st_read(f)
  shp <- shp %>% st_transform(3857)
  # get name to save
  city_name <- gsub(".*data/relative/(.+)_employment_predictions_relative.gpkg.*", "\\1", f)
  shp_name <- paste0(city_name, "_PE")
  # save with new name
  assign(paste(shp_name), shp, envir = .GlobalEnv)
}
# remove the temporary objects
rm(shp)
```

- features (night lights, worldpop, OSM amenities, OSM street intersections)

```{r}
# basic procedure 
input_path <- "data/features/"
files <- list.files(input_path, pattern = "[.]gpkg$", full.names=T)
for (f in files){
  # read in file
  shp <- st_read(f)
  shp <- shp %>% st_transform(3857)
  # get name to save
  city_name <- gsub(".*data/features/(.+)_all_features.gpkg.*", "\\1", f)
  shp_name <- paste0(city_name, "_features")
  # save with new name
  assign(paste(shp_name), shp, envir = .GlobalEnv)
}
# remove the temporary objects
rm(shp)
```

- building height

```{r}
# basic procedure 
input_path_bh <- "data/GHS-BUILT-H_hex/"
files_bh <- list.files(input_path_bh, pattern = "[.]gpkg$", full.names=T)
for (f in files_bh){
  # read in file
  shp <- st_read(f)
  shp <- shp %>% st_transform(3857)
  shp <- shp[-c(2)]
  names(shp)[2] <- "building_height"
  # drop id 
  # get name to save
  city_name <- gsub(".*data/GHS-BUILT-H_hex/(.+)_hex.gpkg.*", "\\1", f)
  # save with new name
  assign(city_name, shp, envir = .GlobalEnv)
}
# remove the temporary objects
rm(shp)
```

# Moran's I

```{r}
# get list of data 
features_dfs <- mget(grep('_features', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

- Night light (from features)

checking for missing data 

```{r}
for(i in 1:length(features_dfs)) {
  nl_df <- features_dfs[[i]] %>% drop_na(nlights)
  feature_name <- names(features_dfs)[i]
  city_name <- gsub("(.+)_features", "\\1", feature_name)
  df_name_NL <- paste0(city_name, "_NL")
  # change name of data
  assign(paste(df_name_NL), nl_df, envir = .GlobalEnv)
}
rm(nl_df)
```


```{r}
moran_i_cluster_NL <- function(df, density_threshold, df_name){
  df_sub <- df[df[["nlights"]] > density_threshold,]
  # create neighbours list based on distance threshold
  nb_list <- df_sub %>% poly2nb(., queen=T)
  # create spatial weights list
  sp_weights <- nb_list %>% nb2listw(., style = "B", zero.policy=TRUE)
  # calculate morans i 
  lmi <- df_sub %>%
    pull("nlights") %>%
    as.vector()%>%
    localmoran(., sp_weights, zero.policy=TRUE, na.action=na.exclude)
  # convert into dataframe
  lmi_df <- as.data.frame(lmi)
  # rename the columns 
  names(lmi_df) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")
  # join to original sub-dataframe
  df_sub <- df_sub %>% mutate(lmi_df)
  # Calculate the means of our two columns, observed and LMI_Ii, and store as new columns 
  df_sub <- df_sub %>%
    mutate(obs_diff = (df_sub[["nlights"]] - mean(df_sub[["nlights"]], na.rm = TRUE))) %>%
    mutate(LMI_diff = (df_sub$LMI_Ii - mean(df_sub$LMI_Ii, na.rm = TRUE)))
  # Generate column with cluster type using values above and significance
  df_sub <- df_sub %>%
    mutate(cluster_type_NL = case_when(obs_diff > 0 & LMI_diff > 0 ~ "High-High", obs_diff < 0 & LMI_diff < 0 ~ "Low-Low", obs_diff < 0 & LMI_diff > 0 ~ "Low-High", obs_diff > 0 & LMI_diff < 0 ~ "High-Low")) %>%
    mutate(significant_NL = ifelse(LMI_sigP < 0.1, "significant", "not significant")) %>%
    mutate(econ_cluster_NL = ifelse(cluster_type_NL == "High-High" & significant_NL == "significant", 1, 0))
  # subset columns to join to the original dataframe
  df_join <- df_sub %>%
    select(grid_id, econ_cluster_NL) %>%
    st_drop_geometry()
  # join to the original dataframe
  df <- df %>% 
    left_join(df_join, by='grid_id') %>%
    select(grid_id, econ_cluster_NL) 
  # replace the NAs that occurred with the join procedure with 0
  df[["econ_cluster_NL"]][is.na(df[["econ_cluster_NL"]])] <- 0
  # drop the geometry
  df_final <- df %>% st_drop_geometry()
  # change name of data
  assign(paste(df_name), df_final, envir = .GlobalEnv)
}
```

```{r}
# get list of data 
nl_dfs <- mget(grep('_NL', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
for(i in 1:length(nl_dfs)) {
  cluster_threshold_NL <- unname(quantile(nl_dfs[[i]][["nlights"]], probs=0.8))
  cluster_name_NL <- names(nl_dfs)[i]
  moran_i_cluster_NL(nl_dfs[[i]], cluster_threshold_NL, paste(cluster_name_NL))
}
```

- World pop (from features)

checking for missing data 

```{r}
for(i in 1:length(features_dfs)) {
  wp_df <- features_dfs[[i]] %>% drop_na(worldpop)
  feature_name <- names(features_dfs)[i]
  city_name <- gsub("(.+)_features", "\\1", feature_name)
  df_name_WP <- paste0(city_name, "_WP")
  # change name of data
  assign(paste(df_name_WP), wp_df, envir = .GlobalEnv)
}
rm(wp_df)
```

```{r}
moran_i_cluster_WP <- function(df, density_threshold, df_name){
  df_sub <- df[df[["worldpop"]] > density_threshold,]
  # create neighbours list based on distance threshold
  nb_list <- df_sub %>% poly2nb(., queen=T)
  # create spatial weights list
  sp_weights <- nb_list %>% nb2listw(., style = "B", zero.policy=TRUE)
  # calculate morans i 
  lmi <- df_sub %>%
    pull("worldpop") %>%
    as.vector()%>%
    localmoran(., sp_weights, zero.policy=TRUE, na.action=na.exclude)
  # convert into dataframe
  lmi_df <- as.data.frame(lmi)
  # rename the columns 
  names(lmi_df) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")
  # join to original sub-dataframe
  df_sub <- df_sub %>% mutate(lmi_df)
  # Calculate the means of our two columns, observed and LMI_Ii, and store as new columns 
  df_sub <- df_sub %>%
    mutate(obs_diff = (df_sub[["worldpop"]] - mean(df_sub[["worldpop"]], na.rm = TRUE))) %>%
    mutate(LMI_diff = (df_sub$LMI_Ii - mean(df_sub$LMI_Ii, na.rm = TRUE)))
  # Generate column with cluster type using values above and significance
  df_sub <- df_sub %>%
    mutate(cluster_type_WP = case_when(obs_diff > 0 & LMI_diff > 0 ~ "High-High", obs_diff < 0 & LMI_diff < 0 ~ "Low-Low", obs_diff < 0 & LMI_diff > 0 ~ "Low-High", obs_diff > 0 & LMI_diff < 0 ~ "High-Low")) %>%
    mutate(significant_WP = ifelse(LMI_sigP < 0.1, "significant", "not significant")) %>%
    mutate(econ_cluster_WP = ifelse(cluster_type_WP == "High-High" & significant_WP == "significant", 1, 0))
  # subset columns to join to the original dataframe
  df_join <- df_sub %>%
    select(grid_id, econ_cluster_WP) %>%
    st_drop_geometry()
  # join to the original dataframe
  df <- df %>% 
    left_join(df_join, by='grid_id') %>%
    select(grid_id, econ_cluster_WP)
  # replace the NAs that occurred with the join procedure with 0
  df[["econ_cluster_WP"]][is.na(df[["econ_cluster_WP"]])] <- 0
  # drop the geometry
  df_final <- df %>% st_drop_geometry()
  # change name of data
  assign(paste(df_name), df_final, envir = .GlobalEnv)
}
```

```{r}
# get list of data 
wp_dfs <- mget(grep('_WP', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
for(i in 1:length(wp_dfs)) {
  cluster_threshold_WP <- unname(quantile(wp_dfs[[i]][["worldpop"]], probs=0.8, na.rm = TRUE))
  cluster_name_WP <- names(wp_dfs)[i]
  moran_i_cluster_WP(wp_dfs[[i]], cluster_threshold_WP, paste(cluster_name_WP))
}
```

- OSM amenities (from features)

```{r}
moran_i_cluster_AM <- function(df, density_threshold, df_name){
  df_sub <- df[df[["log_amenities_all_dens"]] > density_threshold,]
  # create neighbours list based on distance threshold
  nb_list <- df_sub %>% poly2nb(., queen=T)
  # create spatial weights list
  sp_weights <- nb_list %>% nb2listw(., style = "B", zero.policy=TRUE)
  # calculate morans i 
  lmi <- df_sub %>%
    pull("log_amenities_all_dens") %>%
    as.vector()%>%
    localmoran(., sp_weights, zero.policy=TRUE, na.action=na.exclude)
  # convert into dataframe
  lmi_df <- as.data.frame(lmi)
  # rename the columns 
  names(lmi_df) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")
  # join to original sub-dataframe
  df_sub <- df_sub %>% mutate(lmi_df)
  # Calculate the means of our two columns, observed and LMI_Ii, and store as new columns 
  df_sub <- df_sub %>%
    mutate(obs_diff = (df_sub[["log_amenities_all_dens"]] - mean(df_sub[["log_amenities_all_dens"]], na.rm = TRUE))) %>%
    mutate(LMI_diff = (df_sub$LMI_Ii - mean(df_sub$LMI_Ii, na.rm = TRUE)))
  # Generate column with cluster type using values above and significance
  df_sub <- df_sub %>%
    mutate(cluster_type_AM = case_when(obs_diff > 0 & LMI_diff > 0 ~ "High-High", obs_diff < 0 & LMI_diff < 0 ~ "Low-Low", obs_diff < 0 & LMI_diff > 0 ~ "Low-High", obs_diff > 0 & LMI_diff < 0 ~ "High-Low")) %>%
    mutate(significant_AM = ifelse(LMI_sigP < 0.1, "significant", "not significant")) %>%
    mutate(econ_cluster_AM = ifelse(cluster_type_AM == "High-High" & significant_AM == "significant", 1, 0))
  # subset columns to join to the original dataframe
  df_join <- df_sub %>%
    select(grid_id, econ_cluster_AM) %>%
    st_drop_geometry()
  # join to the original dataframe
  df <- df %>% 
    left_join(df_join, by='grid_id') %>%
    select(grid_id, econ_cluster_AM)
  # replace the NAs that occurred with the join procedure with 0
  df[["econ_cluster_AM"]][is.na(df[["econ_cluster_AM"]])] <- 0
  # drop the geometry
  df_final <- df %>% st_drop_geometry()
  # change name of data
  assign(paste(df_name), df_final, envir = .GlobalEnv)
}
```

```{r}
for(i in 1:length(features_dfs)) {
  cluster_threshold_AM <- unname(quantile(features_dfs[[i]][["log_amenities_all_dens"]], probs=0.8))
  feature_name <- names(features_dfs)[i]
  city_name <- gsub("(.+)_features", "\\1", feature_name)
  cluster_name_AM <- paste0(city_name, "_AM")
  moran_i_cluster_AM(features_dfs[[i]], cluster_threshold_AM, paste(cluster_name_AM))
}
```

- OSM street intersections

```{r}
moran_i_cluster_SI <- function(df, density_threshold, df_name){
  df_sub <- df[df[["log_intersections_all_dens"]] > density_threshold,]
  # create neighbours list based on distance threshold
  nb_list <- df_sub %>% poly2nb(., queen=T)
  # create spatial weights list
  sp_weights <- nb_list %>% nb2listw(., style = "B", zero.policy=TRUE)
  # calculate morans i 
  lmi <- df_sub %>%
    pull("log_intersections_all_dens") %>%
    as.vector()%>%
    localmoran(., sp_weights, zero.policy=TRUE, na.action=na.exclude)
  # convert into dataframe
  lmi_df <- as.data.frame(lmi)
  # rename the columns 
  names(lmi_df) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")
  # join to original sub-dataframe
  df_sub <- df_sub %>% mutate(lmi_df)
  # Calculate the means of our two columns, observed and LMI_Ii, and store as new columns 
  df_sub <- df_sub %>%
    mutate(obs_diff = (df_sub[["log_intersections_all_dens"]] - mean(df_sub[["log_intersections_all_dens"]], na.rm = TRUE))) %>%
    mutate(LMI_diff = (df_sub$LMI_Ii - mean(df_sub$LMI_Ii, na.rm = TRUE)))
  # Generate column with cluster type using values above and significance
  df_sub <- df_sub %>%
    mutate(cluster_type_SI = case_when(obs_diff > 0 & LMI_diff > 0 ~ "High-High", obs_diff < 0 & LMI_diff < 0 ~ "Low-Low", obs_diff < 0 & LMI_diff > 0 ~ "Low-High", obs_diff > 0 & LMI_diff < 0 ~ "High-Low")) %>%
    mutate(significant_SI = ifelse(LMI_sigP < 0.1, "significant", "not significant")) %>%
    mutate(econ_cluster_SI = ifelse(cluster_type_SI == "High-High" & significant_SI == "significant", 1, 0))
  # subset columns to join to the original dataframe
  df_join <- df_sub %>%
    select(grid_id, econ_cluster_SI) %>%
    st_drop_geometry()
  # join to the original dataframe
  df <- df %>% 
    left_join(df_join, by='grid_id') %>%
    select(grid_id, econ_cluster_SI)
  # replace the NAs that occurred with the join procedure with 0
  df[["econ_cluster_SI"]][is.na(df[["econ_cluster_SI"]])] <- 0
  # drop the geometry
  df_final <- df %>% st_drop_geometry()
  # change name of data
  assign(paste(df_name), df_final, envir = .GlobalEnv)
}
```

```{r}
for(i in 1:length(features_dfs)) {
  cluster_threshold_SI <- unname(quantile(features_dfs[[i]][["log_intersections_all_dens"]], probs=0.8))
  feature_name <- names(features_dfs)[i]
  city_name <- gsub("(.+)_features", "\\1", feature_name)
  cluster_name_SI <- paste0(city_name, "_SI")
  moran_i_cluster_SI(features_dfs[[i]], cluster_threshold_SI, paste(cluster_name_SI))
}
```

- Predicted employment (relative) 

```{r}
moran_i_cluster_PE <- function(df, density_threshold, df_name){
  df_sub <- df[df[["pred_relative_employment"]] > density_threshold,]
  # create neighbours list based on distance threshold
  nb_list <- df_sub %>% poly2nb(., queen=T)
  # create spatial weights list
  sp_weights <- nb_list %>% nb2listw(., style = "B", zero.policy=TRUE)
  # calculate morans i 
  lmi <- df_sub %>%
    pull("pred_relative_employment") %>%
    as.vector()%>%
    localmoran(., sp_weights, zero.policy=TRUE, na.action=na.exclude)
  # convert into dataframe
  lmi_df <- as.data.frame(lmi)
  # rename the columns 
  names(lmi_df) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")
  # join to original sub-dataframe
  df_sub <- df_sub %>% mutate(lmi_df)
  # Calculate the means of our two columns, observed and LMI_Ii, and store as new columns 
  df_sub <- df_sub %>%
    mutate(obs_diff = (df_sub[["pred_relative_employment"]] - mean(df_sub[["pred_relative_employment"]], na.rm = TRUE))) %>%
    mutate(LMI_diff = (df_sub$LMI_Ii - mean(df_sub$LMI_Ii, na.rm = TRUE)))
  # Generate column with cluster type using values above and significance
  df_sub <- df_sub %>%
    mutate(cluster_type_PE = case_when(obs_diff > 0 & LMI_diff > 0 ~ "High-High", obs_diff < 0 & LMI_diff < 0 ~ "Low-Low", obs_diff < 0 & LMI_diff > 0 ~ "Low-High", obs_diff > 0 & LMI_diff < 0 ~ "High-Low")) %>%
    mutate(significant_PE = ifelse(LMI_sigP < 0.1, "significant", "not significant")) %>%
    mutate(econ_cluster_PE = ifelse(cluster_type_PE == "High-High" & significant_PE == "significant", 1, 0))
  # subset columns to join to the original dataframe
  df_join <- df_sub %>%
    select(grid_id, econ_cluster_PE) %>%
    st_drop_geometry()
  # join to the original dataframe
  df <- df %>% 
    left_join(df_join, by='grid_id') %>%
    select(grid_id, econ_cluster_PE)
  # replace the NAs that occurred with the join procedure with 0
  df[["econ_cluster_PE"]][is.na(df[["econ_cluster_PE"]])] <- 0
  # drop the geometry
  df_final <- df %>% st_drop_geometry()
  # change name of data
  assign(paste(df_name), df_final, envir = .GlobalEnv)
}
```

```{r}
# get list of data 
pe_dfs <- mget(grep('_PE', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
for(i in 1:length(pe_dfs)) {
  cluster_threshold_PE <- unname(quantile(pe_dfs[[i]][["pred_relative_employment"]], probs=0.8))
  cluster_name_PE <- paste0(names(pe_dfs)[i])
  moran_i_cluster_PE(pe_dfs[[i]], cluster_threshold_PE, cluster_name_PE)
}
```

- Building Height

checking for missing data 

```{r}
# get list of data 
bh_dfs <- mget(grep('_BH', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
for(i in 1:length(bh_dfs)) {
  bh_df <- bh_dfs[[i]] %>% drop_na(building_height)
  df_name_bh <- names(bh_dfs)[i]
  # change name of data
  assign(paste(df_name_bh), bh_df, envir = .GlobalEnv)
}
rm(bh_df)
```

```{r}
# get list of data (again)
bh_dfs <- mget(grep('_BH', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
moran_i_cluster_BH <- function(df, density_threshold, df_name){
  df_sub <- df[df[["building_height"]] > density_threshold,] 
  # create neighbours list based on distance threshold
  nb_list <- df_sub %>% poly2nb(., queen=T)
  # create spatial weights list
  sp_weights <- nb_list %>% nb2listw(., style = "B", zero.policy=TRUE)
  # calculate morans i 
  lmi <- df_sub %>%
    pull("building_height") %>%
    as.vector()%>%
    localmoran(., sp_weights, zero.policy=TRUE, na.action=na.exclude)
  # convert into dataframe
  lmi_df <- as.data.frame(lmi)
  # rename the columns 
  names(lmi_df) <- c("LMI_Ii", "LMI_eIi", "LMI_varIi", "LMI_zIi", "LMI_sigP")
  # join to original sub-dataframe
  df_sub <- df_sub %>% mutate(lmi_df)
  # Calculate the means of our two columns, observed and LMI_Ii, and store as new columns 
  df_sub <- df_sub %>%
    mutate(obs_diff = (df_sub[["building_height"]] - mean(df_sub[["building_height"]], na.rm = TRUE))) %>%
    mutate(LMI_diff = (df_sub$LMI_Ii - mean(df_sub$LMI_Ii, na.rm = TRUE)))
  # Generate column with cluster type using values above and significance
  df_sub <- df_sub %>%
    mutate(cluster_type_BH = case_when(obs_diff > 0 & LMI_diff > 0 ~ "High-High", obs_diff < 0 & LMI_diff < 0 ~ "Low-Low", obs_diff < 0 & LMI_diff > 0 ~ "Low-High", obs_diff > 0 & LMI_diff < 0 ~ "High-Low")) %>%
    mutate(significant_BH = ifelse(LMI_sigP < 0.1, "significant", "not significant")) %>%
    mutate(econ_cluster_BH = ifelse(cluster_type_BH == "High-High" & significant_BH == "significant", 1, 0))
  # subset columns to join to the original dataframe
  df_join <- df_sub %>%
    select(grid_id, econ_cluster_BH) %>%
    st_drop_geometry()
  # join to the original dataframe
  df <- df %>% 
    left_join(df_join, by='grid_id') %>%
    select(grid_id, econ_cluster_BH)
  # replace the NAs that occurred with the join procedure with 0
  df[["econ_cluster_BH"]][is.na(df[["econ_cluster_BH"]])] <- 0
  # drop the geometry
  df_final <- df %>% st_drop_geometry()
  # change name of data
  assign(paste(df_name), df_final, envir = .GlobalEnv)
}
```

```{r}
for(i in 1:length(bh_dfs)) {
  cluster_threshold_BH <- unname(quantile(bh_dfs[[i]][["building_height"]], probs=0.8))
  cluster_name_BH <- paste0(names(bh_dfs)[i])
  moran_i_cluster_BH(bh_dfs[[i]], cluster_threshold_BH, paste(cluster_name_BH))
}
```

# Merge the dataframes

```{r}
for(i in 1:length(features_dfs)){
  city_name_df <- names(features_dfs)[i]
  # specify pattern of owin 
  city_pattern <- gsub("(.+)_features.*", "\\1", city_name_df)
  # select matching dfs
  city_df_list <- Filter(function(x) is(x, "data.frame"), mget(ls(pattern = city_pattern)))
  city_df_list <- city_df_list[-3]
  # merge all the dfs in the list 
  df_list <- city_df_list %>% reduce(left_join, by = "grid_id")
  # save
  feature_name <- names(features_dfs)[i]
  city_name <- gsub("(.+)_features", "\\1", feature_name)
  cluster_name <- paste0(city_name, "_cluster")
  assign(cluster_name, df_list, envir = .GlobalEnv)
}
```

```{r}
# get list of data 
cluster_dfs <- mget(grep('_cluster', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
for(i in 1:length(cluster_dfs)){
  city_name_df <- names(cluster_dfs)[i]
  # specify pattern of owin 
  city_pattern <- gsub("(.+)_cluster.*", "\\1", city_name_df)
  city_all_pattern <- paste0(city_pattern, "_features")
  # select matching df
  city_all_list <- Filter(function(x) is(x, "data.frame"), mget(ls(pattern = city_all_pattern)))
  city_all <- city_all_list[[1]]
  city_cluster <- cluster_dfs[[i]]
  # create the final output
  df_cluster_final <- city_all %>%
    left_join(city_cluster, by = "grid_id") %>%
    select(grid_id, econ_cluster_AM, econ_cluster_BH, econ_cluster_NL, econ_cluster_PE, econ_cluster_SI, econ_cluster_WP)
  df_cluster_final_name <- paste0(names(cluster_dfs)[i])
  assign(df_cluster_final_name, st_sf(df_cluster_final), envir = .GlobalEnv)
}
# remove the temporary files
rm(city_all)
rm(city_cluster)
rm(city_df_list)
rm(city_all_list)
rm(df_cluster_final)
```

# Create the outputs

```{r}
# get list of data 
cluster_dfs <- mget(grep('_cluster', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r, eval=FALSE}
# save their outputs
for(i in 1:length(cluster_dfs)) {
## construct a file name and save this data:
  file.name = paste0("data/cluster_2/", names(cluster_dfs)[i], ".gpkg")
  st_write(cluster_dfs[[i]], file.name)
}
```

# Summarise the results 

```{r}
for(i in 1:length(cluster_dfs)){
  cluster_df <- cluster_dfs[[i]]
  # count number of grids
  cluster_df <- cluster_df %>% mutate(n_grids = 1) %>% st_drop_geometry()
  cluster_agg <- cluster_df %>%
    summarise_if(is.numeric, sum, na.rm = TRUE)
  cluster_df_name <- names(cluster_dfs)[i]
  city_name <- gsub("(.+)_cluster.*", "\\1", cluster_df_name)
  row.names(cluster_agg)[1] <- paste0(city_name)
  cluster_agg_name <- paste0(names(cluster_dfs)[i], "_agg")
  assign(cluster_agg_name, cluster_agg, envir = .GlobalEnv)
}
```

```{r}
# get list of data 
cluster_agg_dfs <- mget(grep('_cluster_agg', names(which(unlist(eapply(.GlobalEnv,is.data.frame)))), value = TRUE))
```

```{r}
# append the dataframes together
cluster_results <- bind_rows(cluster_agg_dfs)
```

```{r}
# export the results
write.csv(cluster_results, "data/cluster_results.csv", row.names=TRUE)
```

